\section{Relevant Publications}
\label{sec:conclusions:pub}

The contributions of this dissertation have been presented to the scientific community in a series of international forums, such as journals and conferences.

\subsection{International JCR journals}

The current version of the EAM learning system was published in the following journal:

\begin{itemize}
 \item Gorka Azkune, Aitor Almeida, Diego López-de-Ipiña, Liming Chen. Extending Knowledge-Driven Activity Models through Data-Driven Learning Techniques. Expert Systems with Applications, vol. 42, no. 6, pp. 3115-3128. Pergamon-Elsevier Science LTD. United States. DOI: 10.1016/j.eswa.2014.11.063, ISSN 0957-4174, JCR Impact Factor (2013): 1.965, Q1. April 2015. 
\end{itemize}

The formalisation of the hybrid evaluation methodology and the synthetic dataset generator was published in the following journal:

\begin{itemize}
 \item Gorka Azkune, Aitor Almeida, Diego López-de-Ipiña, Liming Chen. Combining Users' Activity Survey and Simulators to Evaluate Human Activity Recognition Systems. MDPI Sensors, JCR Impact Factor (2013): 2.048, Q1. Accepted, to be published.
\end{itemize}

An up-to-date survey of the state of the art in activity modelling can be found in the following journal:

\begin{itemize}
 \item Gorka Azkune, Aitor Almeida, Diego López-de-Ipiña, Liming Chen. Latest Trends in Human Activity Modelling. Dyna. JCR Impact Factor (2013): 0.200, Q4. Accepted, to be published. 
\end{itemize}

The preliminary work that led to the research developed in the dissertation was published in the following journal:

\begin{itemize}
 \item Gorka Azkune, Pablo Orduña, Xabier Laiseca, Eduardo Castillejo, Diego López-de-Ipiña, Miguel Loitxate and Jon Azpiazu. Semantic Framework for Social Robot Self-Configuration. MDPI Sensors. vol. 13, no. 6, pp. 7004-7020, DOI: 10.3390/s130607004, ISSN 1424-8220, JCR Impact Factor (2013): 2.048, Q1. Basel, Switzerland, May 2013.
\end{itemize}

\subsection{International conferences}

The Semantic Activity Annotation Algorithm ($SA³$) was presented in the following conference:

\begin{itemize}
 \item Gorka Azkune, Aitor Almeida, Diego López-de-Ipiña, Liming Chen. A Knowledge-Driven Tool for Automatic Activity Dataset Annotation. Proceedings of the 7th International Conference Intelligent Systems IEEE IS’2014, September 24-26, 2014, Warsaw, Poland, Volume 1: Mathematical Foundations, Theory, Analyses. ISBN: 978-3-319-11312-8, pp. 593-604.
\end{itemize}

The hybrid evaluation methodology was presented in the following conference:

\begin{itemize}
 \item Gorka Azkune, Aitor Almeida, Diego López-de-Ipiña, Liming Chen. A Hybrid Evaluation Methodology for Human Activity Recognition Systems. In Proceedings of 8th International Conference on Ubiquitous Computing \& Ambient Intelligence, UCAmI 2014, December 2-4, 2014, Belfast, United Kingdom. ISBN: 978-2-319-13101-6, pp. 92-99.
\end{itemize}

Preliminary work which led to the research done in this dissertation was presented in the following conferences:

\begin{itemize}
 \item Gorka Azkune, Pablo Orduña, Xabier Laiseca, Diego López-de-Ipiña, Miguel Loitxate. Semantic Based Self-configuration Approach for Social Robots in Health Care Environments. Ambient Assisted Living and Home Care. Proceedings of 4th International Workconference on Ambient Assisted Living, IWAAL 2012, Vitoria-Gasteiz, Spain, December 3-5, 2012. ISBN: 978-3-642-35394-9, pp. 354-361.
 \item Gorka Azkune, Mikel Astiz, Urko Esnaola, Unai Antero, José Vicente Sogorb, Antonio Alonso. A Navigation System for a High-Speed Professional Cleaning Robot. Towards Autonomous Robotic Systems. Proceedings of 12th Annual Conference, TAROS 2011, Sheffield, UK, August 31 - September 2, 2011. ISBN: 978-3-642-23231-2, pp. 24-35.
\end{itemize}


